Supplementary Report for the article *Functional diversity can facilitate the collapse of an undesirable ecosystem state*

### 2022-09-28 experiments/2_supplement/supplement.Rmd
put growth and inhibition functions section in supplement

@@

(ref:samp-effect1) A simple phenomenological model of the effect of number of strains in an ecosystem on the resilience of that ecosystem, when the resilience of a ecosystem is determined only by the maximum trait value of the strains present in that ecosystem. Each data point is a different ecosystem, with the strains in each ecosystem drawn at random from a pool of 20 strains. Each of the 20 strains was assigned a trait value drawn at random from a uniform distribution with minimum 0 and maximum 1. The blue line is the fit of a generalised additive model and the grey ribbon shows the 1 standard error confidence range.

```{r samp-effect1, echo = FALSE,  include=TRUE, fig.cap='(ref:samp-effect1)', fig.height=5, fig.width=6, fig.align="center"}
if(!("tidyverse" %in% .packages()))
  library(tidyverse)

number_of_species <- 20
species_trait <- tibble(species_name = paste0("Species-", letters[1:20]),
                        trait_value = runif(number_of_species))

@@

fig_div_vs_o2diff_maxtolcomp(all_stab, which_strains,
                              which_wait_time, figure_title)
```

@@

Bush et al. (2018) used the Monod equation for growth, with multiplicative kinetics when two substrates determine growth rate:

g_{max,j} is the maximum specific growth rate of strain $j$

$X$ is the concentration of substrate 1, and $Y$ is the concentration of substrate 2.

K_{j,X} and K_{j,Y} are the half-saturation constants of species $j$ on substrates $X$ and $Y$.

>>>>

The rate of change in population densities is given by the growth rate - death rate [Bush et al 2017](https://www.nature.com/articles/s41467-017-00912-x). Growth rate is the growth rate in absence of any inhibition multiplied by the amount of inhibition.

Growth rate in the absence of inhibition was modeled with the Monod equation with multiplicative kinetics when two substrates determine growth rate:

$g_{max,j}$ is the maximum specific growth rate of strain $j$. $X$ is the concentration of substrate 1, and $Y$ is the concentration of substrate 2. $K_{j,X}$ and $K_{j,Y}$ are the half-saturation constants of species $j$ on substrates $X$ and $Y$.

Inhibition of growth by a substrate was modelled with the Haldane equation:

$h_j(X) = \frac{1}{1+ (X / H_{j,X})}$

$H_{j,X}$ can be thought of as a half-inhibition constant: the concentration of $X$ at which the growth rate of species $j$ is reduced by 50%.

For further information details of the functions used to model system, please see the original publication by Bush *et al.* (2018)

@@

(ref:samp-effect1) A simple phenomenological model of the effect of number of strains in an ecosystem on the resilience of that ecosystem, when the resilience of a ecosystem is linearly proportional to the maximum trait value of the strains present in that ecosystem. Each data point is a different ecosystem, with the strains in each ecosystem drawn at random from a pool of 20 strains. Each of the 20 strains was assigned a trait value drawn at random from a uniform distribution with minimum 0 and maximum 1. The blue line is the fit of a generalised additive model and the grey ribbon shows the 1 standard error confidence range.

```{r samp-effect1, echo = FALSE,  include=TRUE, fig.cap='(ref:samp-effect1)', fig.height=5, fig.width=6, fig.align="center"}
if(!("tidyverse" %in% .packages()))
  library(tidyverse)
set.seed(12345)
number_of_species <- 20
species_trait <- tibble(species_name = paste0("Species-", letters[1:20]),
                        trait_value = runif(number_of_species))

@@

plot_1 <- fig_div_vs_o2diff_maxtolcomp(all_stab, which_strains,
                              which_wait_time, figure_title)
plot_1 +
  xlab("Standardised amount of trait variation used to create nine strains\nof which either all nine are retained, or only the most tolerant is retained.") +
  guides(col=guide_legend(title="Which strains are retained", title.position = "top"))
```

@@
